---
title: "LIBS Calibration Curve"
author: "Komi Messan"
date: "February 26, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, echo=FALSE, message=FALSE, warning=FALSE}
# All packages needed
library(rJava)
library(xlsx)
library(xlsxjars)
library(ggplot2)
library(cowplot)
library(latex2exp)
library(DT) # for datatable
library(pls)  # For Principal Component Regression (PCR) and Partial Least Squares Regression (PLSR)
library(devtools)
library(ggbiplot)
```



```{r, echo=FALSE, warning=FALSE, message=FALSE}
#Read in data
Raw_lead100 <- read.xlsx("Lead_raw_100_ppm.xlsx", sheetName="Sheet1", header = TRUE,stringsAsFactors = F)
Lead_Data <- read.xlsx("Lead_all_averages.xlsx", sheetName="Lead", header = TRUE,stringsAsFactors = F)
CRREL_Soil <- read.xlsx("Lead_all_averages.xlsx", sheetName="CRREL Soils", header = TRUE,stringsAsFactors = F)
Concentration <- c(0,100,1000,10000,100000)
```


# Lead data visualization 

```{r, echo=TRUE, warning=FALSE, message=FALSE, fig.height=25, fig.width=20}
# plot The LIBS Spectra at different concentration

Intensity <- c(Lead_Data$Lead_0_ppm, Lead_Data$Lead_100_ppm, Lead_Data$Lead_1000_ppm, Lead_Data$Lead_10000_ppm, Lead_Data$Lead_100000_ppm)
#Conc <- c(rep(0,12288), rep(100,12288),rep(1000,12288), rep(10000,12288),rep(100000,12288))
Conc <- c(rep("Concentration = 0 ppm",12288),
                          rep("Concentration = 100 ppm",12288),
                          rep("Concentration = 1000 ppm",12288),
                          rep("Concentration = 10000 ppm",12288),
                          rep("Concentration = 100000 ppm",12288))
Wavelength <- rep(Lead_Data$Wavelength,5)

Lead <- data.frame(Wavelength, Intensity,Conc)

head(Lead)


######################################################
black.bold.text1 <- element_text(face = "bold", color = "black",size=26) # x and y axis
black.bold.text2 <- element_text(face = "bold", color = "black",size=22) # title

p<-theme(axis.text.x = element_text(face="bold", color="black", size=22),
         axis.text.y = element_text(face="bold", color="black", size=22),
         title=black.bold.text2,axis.title = black.bold.text1, legend.position = "bottom",
         legend.text = element_text(size=24),strip.text.x = element_text(face="bold",size=22))
##############################################################

ggplot(data = Lead, aes(x = Wavelength, y = Intensity)) +
  geom_line(size = 0.8) + theme_bw() +
  facet_wrap(~Conc,nrow = 5, ncol = 1)+ 
  xlab("Wavelength (nm)") +
  ylab("Intensity (A.U.)") + p

```

# Pinciple Component Analysis of the data

We continue the analysis by doing a principle component analysis (PCA) in order to determine if there exist any significant differences in the elemental composition of the five level of lead concentration (i.e. 0, 100, 1000, 10000, and 100000). Our analysis in R follow description described in [following page](https://www.datacamp.com/community/tutorials/pca-analysis-r).


```{r}
Lead.pca <- prcomp(Lead_Data[,2:6], center = TRUE, scale. = TRUE)


summary(Lead.pca)

ggbiplot(Lead.pca) # visualize the PCA

```


```{r}
ggbiplot(Lead.pca, ellipse = TRUE, groups = Conc) # visualize the PCA
```


Data is now transpose and we add a new column called concentration so we can later do analysis. Analysis will be conducted using partial least square regression from the work of Bjørn-Helge Mevik, 2018 which can be found [here](https://cran.r-project.org/web/packages/pls/vignettes/pls-manual.pdf).


```{r}
mean_data <- t(Lead_Data)
colnames(mean_data) <- mean_data[1,] # make first row to be the column names
mean_data <- mean_data[-1,] # remove the row with wavelength
mean_data <- data.frame(Concentration, mean_data) # add the concentration column

head(mean_data[,1:4])
```






