---
title: "LIBS Calibration Curve"
author: "Komi Messan"
date: "February 26, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, echo=FALSE, message=FALSE, warning=FALSE}
# All packages needed
library(rJava)
library(xlsx)
library(xlsxjars)
library(ggplot2)
library(cowplot)
library(latex2exp)
library(DT) # for datatable
library(pls)  # For Principal Component Regression (PCR) and Partial Least Squares Regression (PLSR)
library(devtools)
library(ggbiplot)
library(openxlsx)
```


# Lead data visualization 

## Exploring negative intensity as a zero

Here we first assume that all negative intensities are zero prior to take the average of the nine LIBS shot.


```{r, echo=FALSE, warning=FALSE, message=FALSE}
#Read in data
Lead_Data <- openxlsx::read.xlsx("Lead_all_averages.xlsx", sheet="Lead", colNames = TRUE)
#CRREL_Soil <- openxlsx::read.xlsx("Lead_all_averages.xlsx", sheet="CRREL Soils", colNames = TRUE)
Concentration <- c(0,100,1000,10000,100000)
```




```{r, echo=TRUE, warning=FALSE, message=FALSE, fig.height=25, fig.width=20}
# plot The LIBS Spectra at different concentration

Intensity <- c(Lead_Data$Lead_0_ppm, Lead_Data$Lead_100_ppm, Lead_Data$Lead_1000_ppm, Lead_Data$Lead_10000_ppm, Lead_Data$Lead_100000_ppm)
#Conc <- c(rep(0,12288), rep(100,12288),rep(1000,12288), rep(10000,12288),rep(100000,12288))
Conc <- c(rep("Concentration = 0 ppm",12288),
                          rep("Concentration = 100 ppm",12288),
                          rep("Concentration = 1000 ppm",12288),
                          rep("Concentration = 10000 ppm",12288),
                          rep("Concentration = 100000 ppm",12288))
Wavelength <- rep(Lead_Data$Wavelength,5)

Lead <- data.frame(Wavelength, Intensity,Conc)

head(Lead)


######################################################
black.bold.text1 <- element_text(face = "bold", color = "black",size=26) # x and y axis
black.bold.text2 <- element_text(face = "bold", color = "black",size=22) # title

p<-theme(axis.text.x = element_text(face="bold", color="black", size=22),
         axis.text.y = element_text(face="bold", color="black", size=22),
         title=black.bold.text2,axis.title = black.bold.text1, legend.position = "bottom",
         legend.text = element_text(size=24),strip.text.x = element_text(face="bold",size=22))
##############################################################

ggplot(data = Lead, aes(x = Wavelength, y = Intensity)) +
  geom_line(size = 0.8) + theme_bw() +
  facet_wrap(~Conc,nrow = 5, ncol = 1)+ 
  xlab("Wavelength (nm)") +
  ylab("Intensity (A.U.)") + p

```

### Pinciple Component Analysis of the data

We continue the analysis by doing a principle component analysis (PCA) in order to determine if there exist any significant differences in the elemental composition of the five level of lead concentration (i.e. 0, 100, 1000, 10000, and 100000). Our analysis in R follow description described in [following page](https://www.datacamp.com/community/tutorials/pca-analysis-r).


```{r, echo=TRUE, warning=FALSE, message=FALSE}
Lead.pca <- prcomp(Lead_Data[,2:6], center = TRUE, scale. = TRUE)


summary(Lead.pca)

ggbiplot(Lead.pca) # visualize the PCA

```


## Exploring negative intensity as missing 

In what follow, we continue analysis by looking at all negative intensity as a missing value.

```{r, echo=FALSE, warning=FALSE, message=FALSE}
#Read in raw data

Raw_lead_0 <- openxlsx::read.xlsx("Lead_raw_data.xlsx", sheet="lead_0_ppm", colNames = TRUE)
Raw_lead_100 <- openxlsx::read.xlsx("Lead_raw_data.xlsx", sheet="lead_100_ppm", colNames = TRUE)
Raw_lead_1000 <- openxlsx::read.xlsx("Lead_raw_data.xlsx", sheet="lead_1000_ppm", colNames = TRUE)
Raw_lead_10000 <- openxlsx::read.xlsx("Lead_raw_data.xlsx", sheet="lead_10000_ppm", colNames = TRUE)
Raw_lead_100000 <- openxlsx::read.xlsx("Lead_raw_data.xlsx", sheet="lead_100000_ppm", colNames = TRUE)
```

 Now we calculate the average of each of the nine by only considering the positive values using the following function
 
```{r, echo=TRUE, warning=FALSE, message=FALSE}

average <- function(Raw_data){
  isPositive <- function(x) x>=0 # function to extract the positive value
  n <- dim(Raw_data)[1] # find the length of the data
  (v <- rep(NA, n)) # create an empty vector of length n
  for (i in 1:n){
    v[i] <- mean(Filter(isPositive, as.numeric(Raw_data[i,-1])))
  }
  v[is.na(v)] <- 0
   return(v)
}

mean_lead_0 <- average(Raw_lead_0)
mean_lead_100 <- average(Raw_lead_100)
mean_lead_1000 <- average(Raw_lead_1000)
mean_lead_10000 <- average(Raw_lead_10000)
mean_lead_100000 <- average(Raw_lead_100000)

Lead_Data_mval <- data.frame(Lead_Data$Wavelength, mean_lead_0, mean_lead_100,
                             mean_lead_1000, mean_lead_10000, 
                             mean_lead_100000)

Intensity_mval <- c(mean_lead_0,mean_lead_100, mean_lead_1000, mean_lead_10000, mean_lead_100000)
Lead_mval <- data.frame(Wavelength, Intensity_mval,Conc) # Wavelength and Conc are in previous block code

```
 
 
 
```{r, echo=TRUE, warning=FALSE, message=FALSE, fig.height=25, fig.width=20}
# plot The LIBS Spectra at different concentration

ggplot(data = Lead_mval, aes(x = Wavelength, y = Intensity_mval)) +
  geom_line(size = 0.8) + theme_bw() +
  facet_wrap(~Conc,nrow = 5, ncol = 1)+
  xlab("Wavelength (nm)") +
  ylab("Intensity (A.U.)") + p

```


 
```{r, echo=TRUE, warning=FALSE, message=FALSE}
Lead_mval.pca <- prcomp(Lead_Data_mval[,2:6], center = TRUE, scale. = TRUE)


ggbiplot(Lead_mval.pca) # visualize the PCA

```





Data is now transpose and we add a new column called concentration so we can later do analysis. Analysis will be conducted using partial least square regression from the work of Bjørn-Helge Mevik, 2018 which can be found [here](https://cran.r-project.org/web/packages/pls/vignettes/pls-manual.pdf).


```{r}
Lead_Data_mval <- data.frame(Lead_Data$Wavelength,mean_lead_0,mean_lead_100,
                             mean_lead_1000, mean_lead_10000,
                             mean_lead_100000)
colnames(Lead_Data_mval) <- c("Wavelength","Lead_0_ppm","Lead_100_ppm",
                              "Lead_1000_ppm","Lead_10000_ppm",
                              "Lead_100000_ppm")

mean_intensity <- t(Lead_Data_mval)
cnames <- mean_intensity[1,]
mean_intensity <- mean_intensity[-1,] # remove the row with wavelength
mean_data_mval <- data.frame(Concentration, mean_intensity=I(mean_intensity)) # add the concentration column but in two different matrix form using I()
colnames(mean_data_mval$mean_intensity) <- cnames # make first row to be the column name


lead.plsr1 <- plsr(Concentration ~ mean_intensity, ncomp = 3, data = mean_data_mval, validation = "LOO")

summary(lead.plsr1)

plot(RMSEP(lead.plsr1), legendpos = "topright") #Root Mean Square Error of the Prediction

 plot(lead.plsr1, ncomp = 2, asp = 1, line = TRUE)


 plot(lead.plsr1, plottype = "scores", comps = 1:3)
 
 plot(lead.plsr1, "loadings", comps = 1:2, legendpos = "topleft",
      labels = "numbers", xlab = "nm")
 abline(h = 0)
```









