---
title: "Multivariate LIBS Calibration Curve"
author: "Komi Messan"
date: "April 5, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r, echo=FALSE, message=FALSE, warning=FALSE}
# All packages needed
library(rJava)
library(xlsx)
library(xlsxjars)
library(ggplot2)
library(cowplot)
library(latex2exp)
library(DT) # for datatable
library(pls)  # For Principal Component Regression (PCR) and Partial Least Squares Regression (PLSR)
library(devtools)
library(ggbiplot)
library(openxlsx)
#library(plotly)
#library(tidyverse)
```


 
 
```{r, echo=FALSE, warning=FALSE, message=FALSE}

# First we create functions to perform average calculation and also the rsquares given appropriate data

average <- function(Raw_data){
  isPositive <- function(x) x>=0 # function to extract the positive value
  se <- function(x) sqrt(var(x)/length(x)) # standard error
  n <- dim(Raw_data)[1] # find the length of the data
  (v_m <- rep(NA, n)) # create an empty vector of length n for the mean
  (v_sd <- rep(NA, n)) # create an empty vector of length n for the sd
  
  for (i in 1:n){
    v_m[i] <- mean(Filter(isPositive, as.numeric(Raw_data[i,-1])))
    v_sd[i] <- sd(Filter(isPositive, as.numeric(Raw_data[i,-1])))
  }
  v_m[is.na(v_m)] <- 0
  v_sd[is.na(v_sd)] <- 0
   return(data.frame(v_m,v_sd))
}


Rvalues <- function(Average_raw_data, concentration){
  
L <- length(Average_raw_data[,1])# length of the intensity data
concentration <- concentration
# create an empty vector for the slope, the intercept and Rsquare values
slope <- numeric(0L)
intercept <- numeric(0L)
Rsquare <- numeric(0L)

# Loop through the entire data and create a linear model where the mean intensity is the dependent variable and the theoretical concentration is the independent variable. 
for (i in 1:L){
  linear_relation <- lm(as.numeric(Average_raw_data[i,2:6])~concentration)
  sum_relation <- summary(linear_relation)
  
  #extract coefficients from linear regression
  slope <- c(slope,linear_relation$coefficients[2])
  intercept <- c(intercept,linear_relation$coefficients[1])
  Rsquare <- c(Rsquare,sum_relation$r.squared)
}  
  
 result <- data.frame(Average_raw_data[,1], slope,intercept, Rsquare)
 colnames(result) <- c("Wavelength","slope","intercept","Rsquare")
 return(result)
}


######## For plotting

black.bold.text1 <- element_text(face = "bold", color = "black",size=18) # x and y axis
black.bold.text2 <- element_text(face = "bold", color = "black",size=18) # title

p<-theme(axis.text.x = element_text(face="bold", color="black", size=16),
         axis.text.y = element_text(face="bold", color="black", size=16),
         title=black.bold.text2,axis.title = black.bold.text1, legend.position = "bottom",
         legend.text = element_text(size=18),strip.text.x = element_text(face="bold",size=18))

```

# Introduction

It is noteworthy to point out that several intensities values are negative due to noises in the data. To process this data, all negative intensities values were treating as missing values in the remaining of the analysis. Thus we have baseline of 0 for the intensities value. The average of each of the nine shots were then taken by only considering the positive values using a pre-defined. Finally, data was transpose and we add a new column called concentration so we can later do analysis. Analysis will be conducted using partial least square regression from the work of Bjørn-Helge Mevik, 2018 which can be found [here](https://cran.r-project.org/web/packages/pls/vignettes/pls-manual.pdf).

# Analysis on Lead data

## Data preprocessing 

```{r, echo=FALSE, warning=FALSE, message=FALSE}
#Read in raw data

Raw_lead_0 <- openxlsx::read.xlsx("Lead_raw_data.xlsx", sheet="lead_0_ppm", colNames = TRUE)
Raw_lead_100 <- openxlsx::read.xlsx("Lead_raw_data.xlsx", sheet="lead_100_ppm", colNames = TRUE)
Raw_lead_1000 <- openxlsx::read.xlsx("Lead_raw_data.xlsx", sheet="lead_1000_ppm", colNames = TRUE)
Raw_lead_10000 <- openxlsx::read.xlsx("Lead_raw_data.xlsx", sheet="lead_10000_ppm", colNames = TRUE)
Raw_lead_100000 <- openxlsx::read.xlsx("Lead_raw_data.xlsx", sheet="lead_100000_ppm", colNames = TRUE)
```


```{r, echo=FALSE, warning=FALSE, message=FALSE}
# Take the average of all nine LIBS shots
mean_lead_0 <- average(Raw_lead_0)$v_m
mean_lead_100 <- average(Raw_lead_100)$v_m
mean_lead_1000 <- average(Raw_lead_1000)$v_m
mean_lead_10000 <- average(Raw_lead_10000)$v_m
mean_lead_100000 <- average(Raw_lead_100000)$v_m

Lead_Data_mval <- data.frame(Raw_lead_0$Wavelength, mean_lead_0,mean_lead_100,
                             mean_lead_1000, mean_lead_10000, 
                             mean_lead_100000)
colnames(Lead_Data_mval) <- c("Wavelength","Lead_0_ppm","Lead_100_ppm",
                              "Lead_1000_ppm","Lead_10000_ppm",
                              "Lead_100000_ppm")

Intensity_mval <- c(mean_lead_0,mean_lead_100, mean_lead_1000, mean_lead_10000, mean_lead_100000)

Conc <- c(rep("Concentration = 0 ppm",12288),
                          rep("Concentration = 100 ppm",12288),
                          rep("Concentration = 1000 ppm",12288),
                          rep("Concentration = 10000 ppm",12288),
                          rep("Concentration = 100000 ppm",12288))
Wavelength <- rep(Raw_lead_0$Wavelength,5)

Lead_mval <- data.frame(Wavelength, Intensity_mval,Conc) # Wavelength and Conc are in previous block code

```
 
 
## Data analysis 

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.height=6, fig.width=14}
# select certain ranges of the wavelength
Lead_mval2 <- Lead_Data_mval[which(Lead_Data_mval$Wavelength>=200 & Lead_Data_mval$Wavelength<=350),] 
pb_concentration <- c(0,100,1000,10000,100000)

pb_mean_intensity <- t(Lead_mval2)
pb_cnames <- pb_mean_intensity[1,]
pb_mean_intensity <- pb_mean_intensity[-1,] # remove the row with wavelength
colnames(pb_mean_intensity) <- pb_cnames # make first row to be the column name
Lead_mval_final <- data.frame(pb_concentration, pb_mean_intensity=I(pb_mean_intensity)) # add the concentration column but in two different matrix form using I()

lead.plsr <- plsr(pb_concentration ~ pb_mean_intensity, ncomp = 4, data = Lead_mval_final)

lead.plsr <- plsr(pb_concentration ~ msc(pb_mean_intensity), ncomp = 4, data = Lead_mval_final,validation = "LOO")



summary(lead.plsr)

lead.par <- par(mfrow=c(1,2))
pb_rmsep <- plot(RMSEP(lead.plsr),xlab="Number of Components",font.axis=4,font.lab=2,main="", type = "o") #Root Mean Square Error of the Prediction
pb_predict_meas <- plot(lead.plsr, ncomp = 2, asp = 1, line = TRUE,
                        xlab="Measured Concentration", 
                        ylab="Predicted Concentration",
                        font.axis=4,font.lab=2, main="") # predicted vs. measured plot
text(1,80000,expression(R^{2} == "0.9047"))
par(lead.par)

```

# Analysis on zinc data

## Data preprocessing

```{r, echo=FALSE, warning=FALSE, message=FALSE}
#Read in raw data

Raw_zinc_0 <- openxlsx::read.xlsx("Zinc_raw_data.xlsx", sheet="zinc_0_ppm", colNames = TRUE)
Raw_zinc_100 <- openxlsx::read.xlsx("Zinc_raw_data.xlsx", sheet="zinc_100_ppm", colNames = TRUE)
Raw_zinc_250 <- openxlsx::read.xlsx("Zinc_raw_data.xlsx", sheet="zinc_250_ppm", colNames = TRUE)
Raw_zinc_500 <- openxlsx::read.xlsx("Zinc_raw_data.xlsx", sheet="zinc_500_ppm", colNames = TRUE)
Raw_zinc_1000 <- openxlsx::read.xlsx("Zinc_raw_data.xlsx", sheet="zinc_1000_ppm", colNames = TRUE)
```


```{r, echo=FALSE, warning=FALSE, message=FALSE}
# Take the average of all nine LIBS shots
mean_zinc_0 <- average(Raw_zinc_0)$v_m
mean_zinc_100 <- average(Raw_zinc_100)$v_m
mean_zinc_250 <- average(Raw_zinc_250)$v_m
mean_zinc_500 <- average(Raw_zinc_500)$v_m
mean_zinc_1000 <- average(Raw_zinc_1000)$v_m

Zinc_Data_mval <- data.frame(Raw_zinc_0$Wavelength, mean_zinc_0,mean_zinc_100,
                             mean_zinc_250, mean_zinc_500, 
                             mean_zinc_1000)

colnames(Zinc_Data_mval) <- c("Wavelength","Zinc_0_ppm","Zinc_100_ppm",
                              "Zinc_250_ppm","Zinc_500_ppm",
                              "Zinc_1000_ppm")

Zinc_Intensity_mval <- c(mean_zinc_0,mean_zinc_100, mean_zinc_250, mean_zinc_500, mean_zinc_1000)

Conc_zinc <- factor(c(rep("Concentration = 0 ppm",12288),
                          rep("Concentration = 100 ppm",12288),
                          rep("Concentration = 250 ppm",12288),
                          rep("Concentration = 500 ppm",12288),
                          rep("Concentration = 1000 ppm",12288)),
                    levels = c("Concentration = 0 ppm",
                               "Concentration = 100 ppm",
                               "Concentration = 250 ppm",
                               "Concentration = 500 ppm",
                               "Concentration = 1000 ppm"))
Wavelength_zinc <- rep(Raw_lead_0$Wavelength,5)

Zinc_mval <- data.frame(Wavelength_zinc, Zinc_Intensity_mval,Conc_zinc) # Wavelength and Conc are in previous block code
colnames(Zinc_mval) <- c("Wavelength","Intensity","Conc")

```

## Data analysis 

```{r, echo=FALSE, warning=FALSE, message=FALSE,  fig.height=6, fig.width=14}

# select certain ranges of the wavelength
Zinc_mval2 <- Zinc_Data_mval[which(Zinc_Data_mval$Wavelength>=200 & Zinc_Data_mval$Wavelength<=350),] 
zn_concentration <- c(0,100,250,500,1000)


zn_mean_intensity <- t(Zinc_mval2)
zn_cnames <- zn_mean_intensity[1,]
zn_mean_intensity <- zn_mean_intensity[-1,] # remove the row with wavelength
colnames(zn_mean_intensity) <- zn_cnames # make first row to be the column name
Zinc_mval_final <- data.frame(zn_concentration, zn_mean_intensity=I(zn_mean_intensity)) # add the concentration column but in two different matrix form using I()



#zinc.plsr <- plsr(zn_concentration ~ zn_mean_intensity, ncomp = 3, data = Zinc_mval_final, validation = "LOO")

zinc.plsr <- plsr(zn_concentration ~ zn_mean_intensity, ncomp = 4, data = Zinc_mval_final)

summary(zinc.plsr)

zinc.par <- par(mfrow=c(1,2))
zn_rmsep <- plot(RMSEP(zinc.plsr),xlab="Number of Components",font.axis=4,font.lab=2,main="", type = "o") #Root Mean Square Error of the Prediction
zn_predict_meas <- plot(zinc.plsr, ncomp = 3, asp = 1, line = TRUE,
                        xlab="Measured Concentration", 
                        ylab="Predicted Concentration",
                        font.axis=4,font.lab=2, main="") # predicted vs. measured plot
text(0,700,expression(R^{2} == "0.8854"))
par(zinc.par)

```


# Analysis on Antimony (Sb) data

## Data preprocessing


```{r, echo=FALSE, warning=FALSE, message=FALSE}
#Read in raw data

Raw_antimony_0 <- openxlsx::read.xlsx("Antimony_raw_data.xlsx", sheet="antimony_0_ppm", colNames = TRUE)
Raw_antimony_100 <- openxlsx::read.xlsx("Antimony_raw_data.xlsx", sheet="antimony_100_ppm", colNames = TRUE)
Raw_antimony_250 <- openxlsx::read.xlsx("Antimony_raw_data.xlsx", sheet="antimony_250_ppm", colNames = TRUE)
Raw_antimony_500 <- openxlsx::read.xlsx("Antimony_raw_data.xlsx", sheet="antimony_500_ppm", colNames = TRUE)
Raw_antimony_1000 <- openxlsx::read.xlsx("Antimony_raw_data.xlsx", sheet="antimony_1000_ppm", colNames = TRUE)
```



```{r, echo=FALSE, warning=FALSE, message=FALSE}
# Take the average of all nine LIBS shots
mean_antimony_0 <- average(Raw_antimony_0)$v_m
mean_antimony_100 <- average(Raw_antimony_100)$v_m
mean_antimony_250 <- average(Raw_antimony_250)$v_m
mean_antimony_500 <- average(Raw_antimony_500)$v_m
mean_antimony_1000 <- average(Raw_antimony_1000)$v_m

Antimony_Data_mval <- data.frame(Raw_antimony_0$Wavelength,mean_antimony_0,mean_antimony_100,mean_antimony_250,mean_antimony_500,mean_antimony_1000)

colnames(Antimony_Data_mval) <- c("Wavelength","Antimony_0_ppm","Antimony_100_ppm",
                              "Antimony_250_ppm","Antimony_500_ppm",
                              "Antimony_1000_ppm")

Antimony_Intensity_mval <- c(mean_antimony_0,mean_antimony_100, mean_antimony_250, mean_antimony_500, mean_antimony_1000)

Conc_antimony <- factor(c(rep("Concentration = 0 ppm",12288),
                          rep("Concentration = 100 ppm",12288),
                          rep("Concentration = 250 ppm",12288),
                          rep("Concentration = 500 ppm",12288),
                          rep("Concentration = 1000 ppm",12288)),
                    levels = c("Concentration = 0 ppm",
                               "Concentration = 100 ppm",
                               "Concentration = 250 ppm",
                               "Concentration = 500 ppm",
                               "Concentration = 1000 ppm"))
Wavelength_antimony <- rep(Raw_antimony_0$Wavelength,5)

Antimony_mval <- data.frame(Wavelength_antimony, Antimony_Intensity_mval,Conc_antimony) # Wavelength and Conc are in previous block code
colnames(Antimony_mval) <- c("Wavelength","Intensity","Conc")

```


## Data analysis 

```{r, echo=FALSE, warning=FALSE, message=FALSE,  fig.height=6, fig.width=14}

# select certain ranges of the wavelength
Antimony_mval2 <- Antimony_Data_mval[which(Antimony_Data_mval$Wavelength>=200 & Antimony_Data_mval$Wavelength<=350),] 
sb_concentration <- c(0,100,250,500,1000)


sb_mean_intensity <- t(Antimony_mval2)
sb_cnames <- sb_mean_intensity[1,]
sb_mean_intensity <- sb_mean_intensity[-1,] # remove the row with wavelength
colnames(sb_mean_intensity) <- sb_cnames # make first row to be the column name
Antimony_mval_final <- data.frame(sb_concentration, sb_mean_intensity=I(sb_mean_intensity)) # add the concentration column but in two different matrix form using I()



# antimony.plsr <- plsr(sb_concentration ~ sb_mean_intensity, ncomp = 3, data = Antimony_mval_final, validation = "LOO")

antimony.plsr <- plsr(sb_concentration ~ sb_mean_intensity, ncomp = 4, data = Antimony_mval_final)

summary(antimony.plsr)

antimony.par <- par(mfrow=c(1,2))
sb_rmsep <- plot(RMSEP(antimony.plsr), xlab="Number of Components",font.axis=4,font.lab=2,main="", type = "o") #Root Mean Square Error of the Prediction
sb_predict_meas <- plot(antimony.plsr, ncomp = 3, asp = 1, line = TRUE,
                        xlab="Measured Concentration", 
                        ylab="Predicted Concentration",
                        font.axis=4,font.lab=2, main="") # predicted vs. measured plot
text(0,900,expression(R^{2} == "0.9973"))
par(antimony.par)

```


# Analysis on Tungsten (W) data

## Handling the negative intensities


```{r, echo=FALSE, warning=FALSE, message=FALSE}
#Read in raw data

Raw_tungsten_0 <- openxlsx::read.xlsx("Tungsten_raw_data.xlsx", sheet="tungsten_0_ppm", colNames = TRUE)
Raw_tungsten_100 <- openxlsx::read.xlsx("Tungsten_raw_data.xlsx", sheet="tungsten_100_ppm", colNames = TRUE)
Raw_tungsten_500 <- openxlsx::read.xlsx("Tungsten_raw_data.xlsx", sheet="tungsten_500_ppm", colNames = TRUE)
Raw_tungsten_1000 <- openxlsx::read.xlsx("Tungsten_raw_data.xlsx", sheet="tungsten_1000_ppm", colNames = TRUE)
Raw_tungsten_10000 <- openxlsx::read.xlsx("Tungsten_raw_data.xlsx", sheet="tungsten_10000_ppm", colNames = TRUE)
```


```{r, echo=FALSE, warning=FALSE, message=FALSE}
# Take the average of all nine LIBS shots
mean_tungsten_0 <- average(Raw_tungsten_0)$v_m
mean_tungsten_100 <- average(Raw_tungsten_100)$v_m
mean_tungsten_500 <- average(Raw_tungsten_500)$v_m
mean_tungsten_1000 <- average(Raw_tungsten_1000)$v_m
mean_tungsten_10000 <- average(Raw_tungsten_10000)$v_m

Tungsten_Data_mval <- data.frame(Raw_tungsten_0$Wavelength,mean_tungsten_0,mean_tungsten_100,mean_tungsten_500,mean_tungsten_1000,mean_tungsten_10000)

colnames(Tungsten_Data_mval) <- c("Wavelength","Tungsten_0_ppm","Tungsten_100_ppm",
                              "Tungsten_500_ppm","Tungsten_1000_ppm",
                              "Tungsten_10000_ppm")

Tungsten_Intensity_mval <- c(mean_tungsten_0,mean_tungsten_100, mean_tungsten_500, mean_tungsten_1000, mean_tungsten_10000)

Conc_tungsten <- factor(c(rep("Concentration = 0 ppm",12288),
                          rep("Concentration = 100 ppm",12288),
                          rep("Concentration = 500 ppm",12288),
                          rep("Concentration = 1000 ppm",12288),
                          rep("Concentration = 10000 ppm",12288)),
                        levels = c("Concentration = 0 ppm",
                                   "Concentration = 100 ppm",
                                   "Concentration = 500 ppm",
                                   "Concentration = 1000 ppm",
                                   "Concentration = 10000 ppm"))
Wavelength_tungsten <- rep(Raw_tungsten_0$Wavelength,5)

Tungsten_mval <- data.frame(Wavelength_tungsten, Tungsten_Intensity_mval,Conc_tungsten) # Wavelength and Conc are in previous block code
colnames(Tungsten_mval) <- c("Wavelength","Intensity","Conc")

```

## Data analysis 

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.height=6, fig.width=14}

# select certain ranges of the wavelength
Tungsten_mval2 <- Tungsten_Data_mval[which(Tungsten_Data_mval$Wavelength>=200 & Tungsten_Data_mval$Wavelength<=350),] 
w_concentration <- c(0,100,500,1000,10000)


w_mean_intensity <- t(Tungsten_mval2)
w_cnames <- w_mean_intensity[1,]
w_mean_intensity <- w_mean_intensity[-1,] # remove the row with wavelength
colnames(w_mean_intensity) <- w_cnames # make first row to be the column name
Tungsten_mval_final <- data.frame(w_concentration, w_mean_intensity=I(w_mean_intensity)) # add the concentration column but in two different matrix form using I()



# tungsten.plsr <- plsr(w_concentration ~ w_mean_intensity, ncomp = 3, data = Tungsten_mval_final, validation = "LOO")

tungsten.plsr <- plsr(w_concentration ~ w_mean_intensity, ncomp = 4, data = Tungsten_mval_final)

summary(tungsten.plsr)

tungsten.par <- par(mfrow=c(1,2))
w_rmsep <- plot(RMSEP(tungsten.plsr), xlab="Number of Components",font.axis=4,font.lab=2,main="", type = "o") #Root Mean Square Error of the Prediction
w_predict_meas <- plot(tungsten.plsr, ncomp = 2, asp = 1, line = TRUE,
                       xlab="Measured Concentration", 
                        ylab="Predicted Concentration",
                        font.axis=4,font.lab=2, main="") # predicted vs. measured plot
text(0,9500,expression(R^{2} == "0.9945"))
par(tungsten.par)

```

