---
title: "Univariate LIBS Calibration curve"
author: "Komi Messan"
date: "March 5, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r, echo=FALSE, message=FALSE, warning=FALSE}
# All packages needed
library(rJava)
library(xlsx)
library(xlsxjars)
library(ggplot2)
library(cowplot)
library(latex2exp)
library(DT) # for datatable
library(pls)  # For Principal Component Regression (PCR) and Partial Least Squares Regression (PLSR)
library(devtools)
library(ggbiplot)
library(openxlsx)
#library(plotly)
#library(tidyverse)
```


 
 
```{r, echo=FALSE, warning=FALSE, message=FALSE}

# First we create functions to perform average calculation and also the rsquares given appropriate data

average <- function(Raw_data){
  isPositive <- function(x) x>=0 # function to extract the positive value
  se <- function(x) sqrt(var(x)/length(x)) # standard error
  n <- dim(Raw_data)[1] # find the length of the data
  (v_m <- rep(NA, n)) # create an empty vector of length n for the mean
  (v_sd <- rep(NA, n)) # create an empty vector of length n for the sd
  
  for (i in 1:n){
    v_m[i] <- mean(Filter(isPositive, as.numeric(Raw_data[i,-1])))
    v_sd[i] <- sd(Filter(isPositive, as.numeric(Raw_data[i,-1])))
  }
  v_m[is.na(v_m)] <- 0
  v_sd[is.na(v_sd)] <- 0
   return(data.frame(v_m,v_sd))
}


Rvalues <- function(Average_raw_data, concentration){
  
L <- length(Average_raw_data[,1])# length of the intensity data
concentration <- concentration
# create an empty vector for the slope, the intercept and Rsquare values
slope <- numeric(0L)
intercept <- numeric(0L)
Rsquare <- numeric(0L)

# Loop through the entire data and create a linear model where the mean intensity is the dependent variable and the theoretical concentration is the independent variable. 
for (i in 1:L){
  linear_relation <- lm(as.numeric(Average_raw_data[i,2:6])~concentration)
  sum_relation <- summary(linear_relation)
  
  #extract coefficients from linear regression
  slope <- c(slope,linear_relation$coefficients[2])
  intercept <- c(intercept,linear_relation$coefficients[1])
  Rsquare <- c(Rsquare,sum_relation$r.squared)
}  
  
 result <- data.frame(Average_raw_data[,1], slope,intercept, Rsquare)
 colnames(result) <- c("Wavelength","slope","intercept","Rsquare")
 return(result)
}


######## For plotting

black.bold.text1 <- element_text(face = "bold", color = "black",size=18) # x and y axis
black.bold.text2 <- element_text(face = "bold", color = "black",size=18) # title

p<-theme(axis.text.x = element_text(face="bold", color="black", size=16),
         axis.text.y = element_text(face="bold", color="black", size=16),
         title=black.bold.text2,axis.title = black.bold.text1, legend.position = "bottom",
         legend.text = element_text(size=18),strip.text.x = element_text(face="bold",size=18))

```


# Analysis on Lead data

## Handling the negative intensities 

It is noteworthy to point out that several intensities values are negative due to noises in the data. To process this data, all negative intensities values were treating as missing values in the remaining of the analysis. Thus we have baseline of 0 for the intensities value.

```{r, echo=FALSE, warning=FALSE, message=FALSE}
#Read in raw data

Raw_lead_0 <- openxlsx::read.xlsx("Lead_raw_data.xlsx", sheet="lead_0_ppm", colNames = TRUE)
Raw_lead_100 <- openxlsx::read.xlsx("Lead_raw_data.xlsx", sheet="lead_100_ppm", colNames = TRUE)
Raw_lead_1000 <- openxlsx::read.xlsx("Lead_raw_data.xlsx", sheet="lead_1000_ppm", colNames = TRUE)
Raw_lead_10000 <- openxlsx::read.xlsx("Lead_raw_data.xlsx", sheet="lead_10000_ppm", colNames = TRUE)
Raw_lead_100000 <- openxlsx::read.xlsx("Lead_raw_data.xlsx", sheet="lead_100000_ppm", colNames = TRUE)
```

The average of each of the nine shots were taken by only considering the positive values using a pre-defined function and also a function to calculate all Rsquare values


```{r, echo=FALSE, warning=FALSE, message=FALSE}
# Take the average of all nine LIBS shots
mean_lead_0 <- average(Raw_lead_0)$v_m
mean_lead_100 <- average(Raw_lead_100)$v_m
mean_lead_1000 <- average(Raw_lead_1000)$v_m
mean_lead_10000 <- average(Raw_lead_10000)$v_m
mean_lead_100000 <- average(Raw_lead_100000)$v_m

Lead_Data_mval <- data.frame(Raw_lead_0$Wavelength, mean_lead_0,mean_lead_100,
                             mean_lead_1000, mean_lead_10000, 
                             mean_lead_100000)

Intensity_mval <- c(mean_lead_0,mean_lead_100, mean_lead_1000, mean_lead_10000, mean_lead_100000)

Conc <- c(rep("Concentration = 0 ppm",12288),
                          rep("Concentration = 100 ppm",12288),
                          rep("Concentration = 1000 ppm",12288),
                          rep("Concentration = 10000 ppm",12288),
                          rep("Concentration = 100000 ppm",12288))
Wavelength <- rep(Raw_lead_0$Wavelength,5)

Lead_mval <- data.frame(Wavelength, Intensity_mval,Conc) # Wavelength and Conc are in previous block code

```
 
 
 
```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.height=10, fig.width=8}
# plot The LIBS Spectra at different concentration

# Lead_mval[which(Lead_mval$Wavelength>280.1 & Lead_mval$Wavelength<280.2 & Lead_mval$Conc=="Concentration = 0 ppm"),] # to check specific values
Lead_mval2 <- Lead_mval[which(Lead_mval$Wavelength>=200 & Lead_mval$Wavelength<350),] # cut the wavelength at 720

#Lead_mval2 <- Lead_mval[which(Lead_mval$Wavelength>=280 & Lead_mval$Wavelength<285),] 
spec_plot_pb <- ggplot(data = Lead_mval2, aes(x = Wavelength, y = Intensity_mval)) +
  geom_line(size = 0.8) + theme_bw() +
  facet_wrap(~Conc,nrow = 5, ncol = 1, scales = "free_y")+
  xlab("Wavelength (nm)") +
  ylab("Intensity (A.U.)") +  scale_x_continuous(breaks=seq(200,350,10)) +
  scale_y_continuous(breaks=seq(0,14000,1000)) + p

spec_plot_pb
```

## Calibration curve for lead at 280.2 nm

From the access to the National Institute of Standards and Technology (NIST) atomic emission spectrum database which can be found here [here](https://physics.nist.gov/PhysRefData/Handbook), the higher intensity of lead (Pb) occurs at 280.1995 which is approximately equal to 280.2 thus we select 280.2 for our univariate analysis. 

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.height=5, fig.width=8}
# calculate the mean and standard error for the univariate analysis
# use "lead strong line" to find the lead info online
#R_pb <- Rvalues(Lead_Data_mval,c(0,100,1000,10000,100000))

# R_pb[which(R_pb$Wavelength>206 & R_pb$Wavelength<207),]
# R_pb[which(R_pb$Rsquare>0.97 & R_pb$Rsquare<0.99 & R_pb$Wavelength<500 & R_zn$slope>0),]

lamb_pb <- 280.249

m_lead_0 <- average(Raw_lead_0[(Raw_lead_0$Wavelength==lamb_pb),])$v_m
m_lead_100 <- average(Raw_lead_100[(Raw_lead_100$Wavelength==lamb_pb),])$v_m
m_lead_1000 <- average(Raw_lead_1000[(Raw_lead_1000$Wavelength==lamb_pb),])$v_m
m_lead_10000 <- average(Raw_lead_10000[(Raw_lead_10000$Wavelength==lamb_pb),])$v_m
m_lead_100000 <- average(Raw_lead_100000[(Raw_lead_100000$Wavelength==lamb_pb),])$v_m


sd_lead_0 <- average(Raw_lead_0[(Raw_lead_0$Wavelength==lamb_pb),])$v_sd
sd_lead_100 <- average(Raw_lead_100[(Raw_lead_100$Wavelength==lamb_pb),])$v_sd
sd_lead_1000 <- average(Raw_lead_1000[(Raw_lead_1000$Wavelength==lamb_pb),])$v_sd
sd_lead_10000 <- average(Raw_lead_10000[(Raw_lead_10000$Wavelength==lamb_pb),])$v_sd
sd_lead_100000 <- average(Raw_lead_100000[(Raw_lead_100000$Wavelength==lamb_pb),])$v_sd

lead_uni_data <- data.frame(c(m_lead_0,m_lead_100,m_lead_1000,
                              m_lead_10000,m_lead_100000),
                            c(sd_lead_0,sd_lead_100,sd_lead_1000,
                              sd_lead_10000,sd_lead_100000),
                            c(1,100,1000,10000,100000))
colnames(lead_uni_data) <- c("Mean_I","sd","Concentration")

## Plotting

summary(lm(Mean_I~Concentration, data = lead_uni_data))

fit_lead <- lm(Mean_I~Concentration, data = lead_uni_data) # Rsquare = 0.6647
 
cali_plot_pb <- ggplot(lead_uni_data, aes(x=Concentration, y=Mean_I)) +
  geom_point(size=2) +
  geom_errorbar(aes(ymin=Mean_I-sd, ymax=Mean_I+sd), colour="black") +
  scale_x_continuous(label= comma, breaks=c(10,1000,100000), trans = "log10")+
  stat_smooth(method = "lm", col = "red", se= FALSE) +
  xlab("Log(Concentration of Pb at 280.2 nm in ppm)") +
  ylab("Intensity (a.u.)") +
  geom_label(x=0.2, y=500, label="R^{2}==0.6647", parse = TRUE) + 
  theme_bw() + p

cali_plot_pb

```


# Analysis on zinc data

## Handling the negative intensities



```{r, echo=FALSE, warning=FALSE, message=FALSE}
#Read in raw data

Raw_zinc_0 <- openxlsx::read.xlsx("Zinc_raw_data.xlsx", sheet="zinc_0_ppm", colNames = TRUE)
Raw_zinc_100 <- openxlsx::read.xlsx("Zinc_raw_data.xlsx", sheet="zinc_100_ppm", colNames = TRUE)
Raw_zinc_250 <- openxlsx::read.xlsx("Zinc_raw_data.xlsx", sheet="zinc_250_ppm", colNames = TRUE)
Raw_zinc_500 <- openxlsx::read.xlsx("Zinc_raw_data.xlsx", sheet="zinc_500_ppm", colNames = TRUE)
Raw_zinc_1000 <- openxlsx::read.xlsx("Zinc_raw_data.xlsx", sheet="zinc_1000_ppm", colNames = TRUE)
```

Now we take the average of all the nine positive shots by considering only the positive intensities


```{r, echo=FALSE, warning=FALSE, message=FALSE}
# Take the average of all nine LIBS shots
mean_zinc_0 <- average(Raw_zinc_0)$v_m
mean_zinc_100 <- average(Raw_zinc_100)$v_m
mean_zinc_250 <- average(Raw_zinc_250)$v_m
mean_zinc_500 <- average(Raw_zinc_500)$v_m
mean_zinc_1000 <- average(Raw_zinc_1000)$v_m

Zinc_Data_mval <- data.frame(Raw_zinc_0$Wavelength, mean_zinc_0,mean_zinc_100,
                             mean_zinc_250, mean_zinc_500, 
                             mean_zinc_1000)

Zinc_Intensity_mval <- c(mean_zinc_0,mean_zinc_100, mean_zinc_250, mean_zinc_500, mean_zinc_1000)

Conc_zinc <- factor(c(rep("Concentration = 0 ppm",12288),
                          rep("Concentration = 100 ppm",12288),
                          rep("Concentration = 250 ppm",12288),
                          rep("Concentration = 500 ppm",12288),
                          rep("Concentration = 1000 ppm",12288)),
                    levels = c("Concentration = 0 ppm",
                               "Concentration = 100 ppm",
                               "Concentration = 250 ppm",
                               "Concentration = 500 ppm",
                               "Concentration = 1000 ppm"))
Wavelength_zinc <- rep(Raw_lead_0$Wavelength,5)

Zinc_mval <- data.frame(Wavelength_zinc, Zinc_Intensity_mval,Conc_zinc) # Wavelength and Conc are in previous block code
colnames(Zinc_mval) <- c("Wavelength","Intensity","Conc")

```




```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.height=10, fig.width=8}

# Zinc_mval[which(Zinc_mval$Wavelength>472 & Zinc_mval$Wavelength<472.8 & Zinc_mval$Conc=="Concentration = 0 ppm"),] # to check specific values

l_w <-200 # lowest value of wavelength to display
u_w <-350 # upper value of wavelength to display
  
Zinc_mval2 <- Zinc_mval[which(Zinc_mval$Wavelength>=l_w & Zinc_mval$Wavelength<u_w),] # cut the wavelength at 720

spec_plot_zn <- ggplot(data = Zinc_mval2, aes(x = Wavelength, y = Intensity))+
  geom_line(size = 0.8) + theme_bw() +
  facet_wrap(~Conc,nrow = 5, ncol = 1, scales = "free_y")+
  xlab("Wavelength (nm)") +
  ylab("Intensity (A.U.)") +  scale_x_continuous(breaks=seq(l_w,u_w,10)) +
  scale_y_continuous(breaks=seq(0,10000,2000)) + p

spec_plot_zn

```


## Calibration curve for zinc at 213.85735 nm

From the access to the National Institute of Standards and Technology (NIST) atomic emission spectrum database, the higher intensity of zinc (Zn) occurs at 213.85735 which is approximately equal to 213.8 thus we select 213.8 for our univariate analysis. 

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.height=5, fig.width=8}
# calculate the mean and standard error for the univariate analysis
#lamb_zn <- 481.017 #rsqure = 0.3559

#lamb_zn <- 472.140

#R_zn <- Rvalues(Zinc_Data_mval,c(0,100,250,500,1000))

# R_zn[which(R_zn$Wavelength>206 & R_zn$Wavelength<207),]
# R_zn[which(R_zn$Rsquare>0.97 & R_zn$Rsquare<0.99 & R_zn$Wavelength<500 & R_zn$slope>0),]

#lamb_zn <- 492.465
#lamb_zn <- 255.475 #rsquare = 0.8
lamb_zn <- 213.886 #rsquare = 0.65

m_zinc_0 <- average(Raw_zinc_0[(Raw_zinc_0$Wavelength==lamb_zn),])$v_m
m_zinc_100 <- average(Raw_zinc_100[(Raw_zinc_100$Wavelength==lamb_zn),])$v_m
m_zinc_250 <- average(Raw_zinc_250[(Raw_zinc_250$Wavelength==lamb_zn),])$v_m
m_zinc_500 <- average(Raw_zinc_500[(Raw_zinc_500$Wavelength==lamb_zn),])$v_m
m_zinc_1000 <- average(Raw_zinc_1000[(Raw_zinc_1000$Wavelength==lamb_zn),])$v_m


sd_zinc_0 <- average(Raw_zinc_0[(Raw_zinc_0$Wavelength==lamb_zn),])$v_sd
sd_zinc_100 <- average(Raw_zinc_100[(Raw_zinc_100$Wavelength==lamb_zn),])$v_sd
sd_zinc_250 <- average(Raw_zinc_250[(Raw_zinc_250$Wavelength==lamb_zn),])$v_sd
sd_zinc_500 <- average(Raw_zinc_500[(Raw_zinc_500$Wavelength==lamb_zn),])$v_sd
sd_zinc_1000 <- average(Raw_zinc_1000[(Raw_zinc_1000$Wavelength==lamb_zn),])$v_sd

zinc_uni_data <- data.frame(c(m_zinc_0,m_zinc_100,m_zinc_250,
                              m_zinc_500,m_zinc_1000),
                            c(sd_zinc_0,sd_zinc_100,sd_zinc_250,
                              sd_zinc_500,sd_zinc_1000),
                            c(0,100,250,500,1000))
colnames(zinc_uni_data) <- c("Mean_I","sd","Concentration")

## Plotting
summary(lm(Mean_I~Concentration, data = zinc_uni_data))

fit_zinc <- lm(Mean_I~Concentration, data = zinc_uni_data) # Rsquare = 0.3459

cali_plot_zn <- ggplot(zinc_uni_data, aes(x=Concentration, y=Mean_I)) +
  geom_point(size=2) +
  geom_errorbar(aes(ymin=Mean_I-sd, ymax=Mean_I+sd), colour="black") +
  scale_x_continuous(label= comma) +
  stat_smooth(method = "lm", col = "red", se= FALSE) +
  xlab("Concentration of Zn at 492.4 nm in ppm") +
  ylab("Intensity (a.u.)") +
  geom_label(x=10, y=60, label="R^2 == 0.6549", parse = TRUE) + 
  theme_bw() + p

cali_plot_zn

```




# Analysis on Antimony (Sb) data

## Handling the negative intensities


```{r, echo=FALSE, warning=FALSE, message=FALSE}
#Read in raw data

Raw_antimony_0 <- openxlsx::read.xlsx("Antimony_raw_data.xlsx", sheet="antimony_0_ppm", colNames = TRUE)
Raw_antimony_100 <- openxlsx::read.xlsx("Antimony_raw_data.xlsx", sheet="antimony_100_ppm", colNames = TRUE)
Raw_antimony_250 <- openxlsx::read.xlsx("Antimony_raw_data.xlsx", sheet="antimony_250_ppm", colNames = TRUE)
Raw_antimony_500 <- openxlsx::read.xlsx("Antimony_raw_data.xlsx", sheet="antimony_500_ppm", colNames = TRUE)
Raw_antimony_1000 <- openxlsx::read.xlsx("Antimony_raw_data.xlsx", sheet="antimony_1000_ppm", colNames = TRUE)
```

Now we take the average of all the nine positive shots by considering only the positive intensities


```{r, echo=FALSE, warning=FALSE, message=FALSE}
# Take the average of all nine LIBS shots
mean_antimony_0 <- average(Raw_antimony_0)$v_m
mean_antimony_100 <- average(Raw_antimony_100)$v_m
mean_antimony_250 <- average(Raw_antimony_250)$v_m
mean_antimony_500 <- average(Raw_antimony_500)$v_m
mean_antimony_1000 <- average(Raw_antimony_1000)$v_m

Antimony_Data_mval <- data.frame(Raw_antimony_0$Wavelength,mean_antimony_0,mean_antimony_100,mean_antimony_250,mean_antimony_500,mean_antimony_1000)

Antimony_Intensity_mval <- c(mean_antimony_0,mean_antimony_100, mean_antimony_250, mean_antimony_500, mean_antimony_1000)

Conc_antimony <- factor(c(rep("Concentration = 0 ppm",12288),
                          rep("Concentration = 100 ppm",12288),
                          rep("Concentration = 250 ppm",12288),
                          rep("Concentration = 500 ppm",12288),
                          rep("Concentration = 1000 ppm",12288)),
                    levels = c("Concentration = 0 ppm",
                               "Concentration = 100 ppm",
                               "Concentration = 250 ppm",
                               "Concentration = 500 ppm",
                               "Concentration = 1000 ppm"))
Wavelength_antimony <- rep(Raw_antimony_0$Wavelength,5)

Antimony_mval <- data.frame(Wavelength_antimony, Antimony_Intensity_mval,Conc_antimony) # Wavelength and Conc are in previous block code
colnames(Antimony_mval) <- c("Wavelength","Intensity","Conc")

```




```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.height=10, fig.width=8}

# Antimony_mval[which(Antimony_mval$Wavelength>231 & Antimony_mval$Wavelength<231.4 & Antimony_mval$Conc=="Concentration = 0 ppm"),] # to check specific values

l_w <-200 # lowest value of wavelength to display
u_w <-350 # upper value of wavelength to display
  
Antimony_mval2 <- Antimony_mval[which(Antimony_mval$Wavelength>=l_w & Antimony_mval$Wavelength<u_w),] # cut the wavelength at 720

spec_plot_sb <- ggplot(data = Antimony_mval2, aes(x = Wavelength, y = Intensity))+
  geom_line(size = 0.8) + theme_bw() +
  facet_wrap(~Conc,nrow = 5, ncol = 1, scales = "free_y")+
  xlab("Wavelength (nm)") +
  ylab("Intensity (A.U.)") +  scale_x_continuous(breaks=seq(l_w,u_w,10)) +
  scale_y_continuous(breaks=seq(0,10000,2000)) + p

spec_plot_sb

```


## Calibration curve for antimony at 276.9931 nm

From the access to the National Institute of Standards and Technology (NIST) atomic emission spectrum database, the higher intensity of Antimony (Sb) occurs at 276.9931 which is approximately equal to 276.9 thus we select 276.9 for our univariate analysis. 

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.height=5, fig.width=8}
# calculate the mean and standard error for the univariate analysis

#lamb_sb <- 231.154 # rsquare = 0.5355

lamb_sb <- 276.956 # rsquare = 0.8994

m_antimony_0 <- average(Raw_antimony_0[(Raw_antimony_0$Wavelength==lamb_sb),])$v_m
m_antimony_100 <- average(Raw_antimony_100[(Raw_antimony_100$Wavelength==lamb_sb),])$v_m
m_antimony_250 <- average(Raw_antimony_250[(Raw_antimony_250$Wavelength==lamb_sb),])$v_m
m_antimony_500 <- average(Raw_antimony_500[(Raw_antimony_500$Wavelength==lamb_sb),])$v_m
m_antimony_1000 <- average(Raw_antimony_1000[(Raw_antimony_1000$Wavelength==lamb_sb),])$v_m


sd_antimony_0 <- average(Raw_antimony_0[(Raw_antimony_0$Wavelength==lamb_sb),])$v_sd
sd_antimony_100 <- average(Raw_antimony_100[(Raw_antimony_100$Wavelength==lamb_sb),])$v_sd
sd_antimony_250 <- average(Raw_antimony_250[(Raw_antimony_250$Wavelength==lamb_sb),])$v_sd
sd_antimony_500 <- average(Raw_antimony_500[(Raw_antimony_500$Wavelength==lamb_sb),])$v_sd
sd_antimony_1000 <- average(Raw_antimony_1000[(Raw_antimony_1000$Wavelength==lamb_sb),])$v_sd

antimony_uni_data <- data.frame(c(m_antimony_0,m_antimony_100,m_antimony_250,
                              m_antimony_500,m_antimony_1000),
                            c(sd_antimony_0,sd_antimony_100,sd_antimony_250,
                              sd_antimony_500,sd_antimony_1000),
                            c(0,100,250,500,1000))
colnames(antimony_uni_data) <- c("Mean_I","sd","Concentration")

## Plotting
summary(lm(Mean_I~Concentration, data = antimony_uni_data))

fit_antimony <- lm(Mean_I~Concentration, data = antimony_uni_data) # Rsquare = 0.0574

cali_plot_sb <- ggplot(antimony_uni_data, aes(x=Concentration, y=Mean_I)) +
  geom_point(size=2) +
  geom_errorbar(aes(ymin=Mean_I-sd, ymax=Mean_I+sd), colour="black") +
  scale_x_continuous(label= comma) +
  stat_smooth(method = "lm", col = "red", se= FALSE) +
  xlab("Concentration of Sb at 276.9 nm in ppm") +
  ylab("Intensity (a.u.)") +
  geom_label(x=10, y=75, label="R^2 == 0.8994", parse = TRUE) + 
  theme_bw() + p

cali_plot_sb

```


# Analysis on Tungsten (W) data

## Handling the negative intensities


```{r, echo=FALSE, warning=FALSE, message=FALSE}
#Read in raw data

Raw_tungsten_0 <- openxlsx::read.xlsx("Tungsten_raw_data.xlsx", sheet="tungsten_0_ppm", colNames = TRUE)
Raw_tungsten_100 <- openxlsx::read.xlsx("Tungsten_raw_data.xlsx", sheet="tungsten_100_ppm", colNames = TRUE)
Raw_tungsten_500 <- openxlsx::read.xlsx("Tungsten_raw_data.xlsx", sheet="tungsten_500_ppm", colNames = TRUE)
Raw_tungsten_1000 <- openxlsx::read.xlsx("Tungsten_raw_data.xlsx", sheet="tungsten_1000_ppm", colNames = TRUE)
Raw_tungsten_10000 <- openxlsx::read.xlsx("Tungsten_raw_data.xlsx", sheet="tungsten_10000_ppm", colNames = TRUE)
```

Now we take the average of all the nine positive shots by considering only the positive intensities


```{r, echo=FALSE, warning=FALSE, message=FALSE}
# Take the average of all nine LIBS shots
mean_tungsten_0 <- average(Raw_tungsten_0)$v_m
mean_tungsten_100 <- average(Raw_tungsten_100)$v_m
mean_tungsten_500 <- average(Raw_tungsten_500)$v_m
mean_tungsten_1000 <- average(Raw_tungsten_1000)$v_m
mean_tungsten_10000 <- average(Raw_tungsten_10000)$v_m

Tungsten_Data_mval <- data.frame(Raw_tungsten_0$Wavelength,mean_tungsten_0,mean_tungsten_100,mean_tungsten_500,mean_tungsten_1000,mean_tungsten_10000)

Tungsten_Intensity_mval <- c(mean_tungsten_0,mean_tungsten_100, mean_tungsten_500, mean_tungsten_1000, mean_tungsten_10000)

Conc_tungsten <- factor(c(rep("Concentration = 0 ppm",12288),
                          rep("Concentration = 100 ppm",12288),
                          rep("Concentration = 500 ppm",12288),
                          rep("Concentration = 1000 ppm",12288),
                          rep("Concentration = 10000 ppm",12288)),
                        levels = c("Concentration = 0 ppm",
                                   "Concentration = 100 ppm",
                                   "Concentration = 500 ppm",
                                   "Concentration = 1000 ppm",
                                   "Concentration = 10000 ppm"))
Wavelength_tungsten <- rep(Raw_tungsten_0$Wavelength,5)

Tungsten_mval <- data.frame(Wavelength_tungsten, Tungsten_Intensity_mval,Conc_tungsten) # Wavelength and Conc are in previous block code
colnames(Tungsten_mval) <- c("Wavelength","Intensity","Conc")

```



```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.height=10, fig.width=8}

# Tungsten_mval[which(Tungsten_mval$Wavelength>429 & Tungsten_mval$Wavelength<430 & Tungsten_mval$Conc=="Concentration = 0 ppm"),] # to check specific values

l_w <-200 # lowest value of wavelength to display
u_w <-350 # upper value of wavelength to display

Tungsten_mval2 <- Tungsten_mval[which(Tungsten_mval$Wavelength>=l_w & Tungsten_mval$Wavelength<u_w),] # cut the wavelength at 720

spec_plot_w <- ggplot(data = Tungsten_mval2, aes(x = Wavelength, y = Intensity))+
  geom_line(size = 0.8) + theme_bw() +
  facet_wrap(~Conc,nrow = 5, ncol = 1, scales = "free_y")+
  xlab("Wavelength (nm)") +
  ylab("Intensity (A.U.)") +  scale_x_continuous(breaks=seq(l_w,u_w,10)) +
  scale_y_continuous(breaks=seq(0,20000,2000)) + p

spec_plot_w

```


## Calibration curve for tungsten at 212.1576 nm

From the access to the National Institute of Standards and Technology (NIST) atomic emission spectrum database, the higher intensity of Tungsten (W) occurs at 212.1576 which is approximately equal to 212.1 thus we select 212.1 nm for our univariate analysis. 

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.height=5, fig.width=8}
# calculate the mean and standard error for the univariate analysis

#lamb_sb <- 231.154 # rsquare = 0.5355

#R_w <- Rvalues(Tungsten_Data_mval,c(0,100,500,1000,10000))

#R_w[which(R_w$Rsquare>0.9 & R_w$Rsquare<0.92 & R_w$Wavelength<530),]

#lamb_w <- 405.245 # rsquare = 0.8994

lamb_w <- 212.052

m_tungsten_0 <- average(Raw_tungsten_0[(Raw_tungsten_0$Wavelength==lamb_w),])$v_m
m_tungsten_100 <- average(Raw_tungsten_100[(Raw_tungsten_100$Wavelength==lamb_w),])$v_m
m_tungsten_500 <- average(Raw_tungsten_500[(Raw_tungsten_500$Wavelength==lamb_w),])$v_m
m_tungsten_1000 <- average(Raw_tungsten_1000[(Raw_tungsten_1000$Wavelength==lamb_w),])$v_m
m_tungsten_10000 <- average(Raw_tungsten_10000[(Raw_tungsten_10000$Wavelength==lamb_w),])$v_m


sd_tungsten_0 <- average(Raw_tungsten_0[(Raw_tungsten_0$Wavelength==lamb_w),])$v_sd
sd_tungsten_100 <- average(Raw_tungsten_100[(Raw_tungsten_100$Wavelength==lamb_w),])$v_sd
sd_tungsten_500 <- average(Raw_tungsten_500[(Raw_tungsten_500$Wavelength==lamb_w),])$v_sd
sd_tungsten_1000 <- average(Raw_tungsten_1000[(Raw_tungsten_1000$Wavelength==lamb_w),])$v_sd
sd_tungsten_10000 <- average(Raw_tungsten_10000[(Raw_tungsten_10000$Wavelength==lamb_w),])$v_sd

tungsten_uni_data <- data.frame(c(m_tungsten_0,m_tungsten_100,m_tungsten_500,
                                  m_tungsten_1000,m_tungsten_10000),
                                c(sd_tungsten_0,sd_tungsten_100,sd_tungsten_500,
                                  sd_tungsten_1000,sd_tungsten_10000),
                                c(1,100,500,1000,10000))
colnames(tungsten_uni_data) <- c("Mean_I","sd","Concentration")

## Plotting
summary(lm(Mean_I~Concentration, data = tungsten_uni_data))

fit_tungsten <- lm(Mean_I~Concentration, data = tungsten_uni_data) # Rsquare = 0.0574

cali_plot_w <- ggplot(tungsten_uni_data, aes(x=Concentration, y=Mean_I)) +
  geom_point(size=2) +
  geom_errorbar(aes(ymin=Mean_I-sd, ymax=Mean_I+sd), colour="black") +
  scale_x_continuous(label= comma, trans = "log10") +
  stat_smooth(method = "lm", col = "red", se= FALSE) +
  xlab("Log(Concentration of W at 212.1 nm in ppm)") +
  ylab("Intensity (a.u.)") +
  geom_label(x=0.1, y=75, label="R^2 == 0.8043", parse = TRUE) + 
  theme_bw() + p

cali_plot_w

```


